---
title: "Machine Learning Project Writeup"
output: html_document
---

###Summary
The purpose of this project is to use data at https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv to build a model and predict the data available at https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv.

###Reading and Cleaning Data
```{r}
#Reading Data
data<-read.csv("pml-training.csv",sep=",")

library(caret)

#Cleaning Data
data1<-data[,7:ncol(data)] #Drop the first 6 columns since I suspect they don't help. Will add them
                           #back if model doesn't work out
data1[(data1=="")]<-NA #Populate empty cells with NA
data1<-data1[,colSums(is.na(data1))==0]  #Removes columns with any NAs
```

###Creating the Model
```{r}
if (!file.exists("rfModel.save")) {
  set.seed(125)
  inTrain<-createDataPartition(y=data1$classe,p=.6,list=FALSE)
  training<-data1[inTrain,]
  testing<-data1[-inTrain,]
  
  rfModel <- train(classe ~ ., data=training, method="rf",prox=TRUE)
  save(rfModel, file="rfModel.save")
} else {
  load("rfModel.save")
}
```
###Out of Sample Error
To get the out of sample error, let's first find the error rate for the test set partitioned from the training data.
```{r}
testing<-data1[-inTrain,]

predict1<-testing$classe==predict(rfModel, testing)
error1<-1-length(predict1[predict1==TRUE])/length(predict1)
error1
```
With this value being so low, I took two more random samples and used the same model.
```{r}
set.seed(1011) #randomly chosen
inTrain2<-createDataPartition(y=data1$classe,p=.4,list=FALSE)
testing2<-data1[inTrain2,]
predict2<-testing2$classe==predict(rfModel, testing2)
error2<-1-(length(predict2[predict2==TRUE])/length(predict2))
error2

set.seed(42) #randomly chosen
inTrain3<-createDataPartition(y=data1$classe,p=.4,list=FALSE)
testing3<-data1[inTrain3,]
predict3<-testing3$classe==predict(rfModel, testing3)
error3<-1-(length(predict3[predict3==TRUE])/length(predict3))
error3
```
The mean out of error sample rate is computed as:
```{r}
errort<-c(error1,error2,error3)
mean(errort)
```